---
title: "Time Series Modelling on GHG"
output: 
  html_document:
        toc: yes
        toc_float: yes
        theme: cosmo
        highlight: espresso
        code_folding: hide
---

```{r echo=FALSE}
time <- format(Sys.time(),"%a %b %d %X %Y")
```

The current time is `r time`.

# 1.0 Data Preparation
## 1.0.1 Load required libraries

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(forecast)
library(fpp2)
library(dplyr)
library(ggplot2)
```

## 1.0.2 Load dataset

```{r}
data <- read.csv("https://raw.githubusercontent.com/data-cracker/datasets/main/MasterDataSet.csv")
glimpse(data)
summary(data)
```

## 1.0.3 Calculate total ghg over total gdp

```{r}
# Sum Up all GHG and GDP of each sector then divide by each other
data$ghg_intensity <- rowSums(data[, 25:34])/rowSums(data[, 2:8])
head(data)
```

# 2.0 Modelling
## 2.0.1 Multivariate Regression

```{r message=FALSE, warning=FALSE}
# Our response variable to be predicted will be ghg_intensity
predictor_variables <- data[,2:23]
response_variable <- data.frame(data[,ncol(data)])
names(response_variable) <- "ghg_intensity"

glimpse(predictor_variables)
head(response_variable)

# Compute the correlation matrix
correlation <- cor(predictor_variables, response_variable)
correlation

highly_correlated <- rownames(correlation)[correlation > 0.7 | correlation < -0.7]
highly_correlated

# Fit a multivariate regression model using selected features
selected_features <- highly_correlated
print(selected_features)

response_variable <- "ghg_intensity"
lm_model <- lm(as.formula(paste(response_variable, "~", paste(selected_features, collapse = " + "))), data = data)

# Print the summary of the model
summary(lm_model)

# Checking if data is overfit by cross validation
# Load required libraries
# install.packages("caret")
library(caret)

# Set up cross-validation
set.seed(123)  # for reproducibility
cv <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

# Train the model using cross-validation
lm_model_cv <- train(
  as.formula(paste(response_variable, "~", paste(selected_features, collapse = " + "))),
  data = data,
  method = "lm",
  trControl = cv
)

# Print cross-validated results
print(lm_model_cv)

residuals <- resid(lm_model)

# Plot Fitted values vs Actual values
plot(lm_model$fitted.values, data$ghg_intensity,
     main = "Fitted Values vs Actual Values",
     xlab = "Fitted Values", ylab = "Actual Values",
     col = "blue", pch = 20)

# Add a line to show where fitted values equal actual values
abline(0, 1, col = "red")

# Add points with an offset for visibility
points(lm_model$fitted.values, data$ghg_intensity + residuals,
       col = "green", pch = 4)

```

## 2.0.2 Model Evaluation - Multivariate Regression

```{r}
# Predict the response variable for the entire dataset
predictions <- predict(lm_model, newdata = data[selected_features])

# Evaluate metrics
mse <- mean((data$ghg_intensity - predictions)^2)
rmse <- sqrt(mse)
mae <- mean(abs(data$ghg_intensity - predictions))
r_squared <- 1 - (sum((data$ghg_intensity - predictions)^2) / sum((data$ghg_intensity - mean(data$ghg_intensity))^2))

# Print the metrics
cat("Mean Squared Error (MSE):", mse, "\n")
cat("Root Mean Squared Error (RMSE):", rmse, "\n")
cat("Mean Absolute Error (MAE):", mae, "\n")
cat("R-squared (R²):", r_squared, "\n")

# Print a comparison between actual and predicted values
comparison <- data.frame(Actual = data$ghg_intensity, Predicted = predictions)
print(comparison)
```

## 2.0.3 Random Forest

```{r}
# Install and load the necessary library
# install.packages("randomForest")
library(randomForest)

# Fit a random forest model using selected features
selected_features <- highly_correlated
selected_features

response_variable <- "ghg_intensity"

rf_formula <- as.formula(paste(response_variable, "~", paste(selected_features, collapse = " + ")))
rf_model <- randomForest(rf_formula, data = data)

# Perform cross-validation
cv_results <- randomForest::randomForest(rf_formula, data = data, nfold = 5)
cv_results

# Print the importance of variables
print(importance(rf_model))

# Print the summary of the model
print(rf_model)
```

## 2.0.4 Model Evaluation - Random Forest

```{r}
# Predict the response variable for the entire dataset
predictions <- predict(rf_model, newdata = data[selected_features])

# Evaluate metrics
mse <- mean((data$ghg_intensity - predictions)^2)
rmse <- sqrt(mse)
mae <- mean(abs(data$ghg_intensity - predictions))
r_squared <- 1 - (sum((data$ghg_intensity - predictions)^2) / sum((data$ghg_intensity - mean(data$ghg_intensity))^2))

# Print the metrics
cat("Mean Squared Error (MSE):", mse, "\n")
cat("Root Mean Squared Error (RMSE):", rmse, "\n")
cat("Mean Absolute Error (MAE):", mae, "\n")
cat("R-squared (R²):", r_squared, "\n")

# Print a comparison between actual and predicted values
comparison <- data.frame(Actual = data$ghg_intensity, Predicted = predictions)
print(comparison)
```

## 2.0.5 ARIMA Model Time Series Forecasting

The ARIMA model is represented as ARIMA(p, d, q), where:

'p' is the order of the autoregressive component.
'd' is the order of differencing needed for stationarity.
'q' is the order of the moving average component.

```{r}
forecast_list <- list()

for (feature in selected_features) {
    ts_data <- ts(data[[feature]], start = min(data$Year), frequency = 1)

    # Fit an ARIMA model to the training data
    fit <- auto.arima(ts_data)

    # Make forecasts on the test data
    forecast_values <- forecast(fit, h = 10)

    forecast_list[[feature]] <- forecast_values$mean
}

forecast_features <- data.frame(forecast_list)
forecast_features
head(forecast_features)

# Use the trained regression model to make predictions for the next ten years
print("Forecasted values for GHG intensity by Multivariate Regression Model")
predictions_lm <- predict(lm_model, newdata = forecast_features)
predictions_lm

# Use the trained random forest model to make predictions for the next ten years
print("Forecasted values for GHG intensity by Random Forest Model")
predictions_rf <- predict(rf_model, newdata = forecast_features)
predictions_rf

# Get the forecasted GHG intensity in year 2030 for both models
years <- 2021:2030

for (i in 1:length(years)) {
  print(paste("Forecasted GHG intensity in year", years[i], "by Regression Model:", predictions_lm[i]))
}

for (i in 1:length(years)) {
  print(paste("Forecasted GHG intensity in year", years[i], "by Random Forest Model:", predictions_rf[i]))
}
```
